This chapter is aimed at the impatient reader who wants a working system as
quickly as possible. The technical detail is deliberately kept at a minimum so,
while you will hopefully end up with something that works, you will not
necessarily understand how it all fits together. For that, please read the
remainder of this guide.

\section{Set Up Your Environment}
We suggest you try this on a 64~bit operating system, as that is better suited
for running \Mimir{}. A 32~bit system would also work, but the maximum sizes for
the indexes wold be limited.

In order to build and run a \Mimir{} server you will need the following pieces
of software installed on your system:
\begin{description}
  \item[Java Development Kit] If you don't have one, you can download one from
  Oracle\footnote{\url{http://www.oracle.com/technetwork/java/javase/downloads/index.html}}.
  Make sure you get the JDK and not the Java Runtime Environment (JRE), as that
  would not be suitable. Once installed, make sure your \verb!JAVA_HOME!
  environment variable points to the location where the JDK was installed.  Make
  sure that the \verb!$JAVA_HOME/bin! location is on your \verb!PATH!.
  \item[Apache ANT] version 1.8.1 or later. You can download it from
  \url{http://ant.apache.org/}. Once installed, make sure your \verb!ANT_HOME!
  environment variable points to the top-level directory of your installation.
  \item[Grails] version 1.3.7. You can download this from
  \url{http://grails.org}. Once installed, make sure your \verb!GRAILS_HOME!
  environment variable points to the top-level directory of your installation.
  \item[Working Internet Connection] The next step, described below, is the
  building of the \Mimir{} library. This starts by automatically downloading all
  the required dependencies, so it requires a working Internet connection. Once
  the software is built, it can work without an remote connection.
  \item[GATE Developer] \Mimir{} is an indexer for GATE Documents. The simplest
  way of generating some GATE documents to be indexed is by using the GATE
  Developer tool\footnote{GATE Developer is available at
  \url{http://gate.ac.uk/download/}. Usage of GATE Developer is beyond the scope
  of this document, so we assume you have a basic understanding of how to use
  it. If not, a good place to start is the tutorials page at
  \url{http://gate.ac.uk/demos/developer-videos/}.}. 
\end{description}
%
\section{Build and Run a \Mimir{} Web Application}
%
After all the prerequisites are installed, we can move to building a \Mimir{}
application. For the purposes of this demo, we will build the {\tt mimir-demo}
application. In a real deployment you may find the alternative application ({\tt
mimir-cloud}) more suitable.

The following steps will help you build the {\tt mimir-demo} application.
Commands that you have to execute are formatted in a distinctive font
\cmd{like this}.
\begin{enumerate}
  \item {\bf Download the \Mimir{} sources}, if you do not already have a copy.
  Pick the appropriate version from one of the tag directories listed in the
  {\tt tags} directory at
  \url{http://gate.svn.sourceforge.net/viewvc/gate/mimir/}, then click on the
  {\em Download GNU tarball} link. If you want to use the latest development
  version (which could be unstable!), use the {\tt trunk} directory instead of a
  tag. Unpack the downloaded file to a location on your computer. If using
  Windows we recommend not using the popular Winzip utility, as that sometimes
  mangles the file names. 7-Zip\footnote{\url{http://www.7-zip.org/}} and the
  Cygwin ``tar'' utility are known to work correctly in this respect, and other
  free archiving tools are available that support the {\tt .tar.gz} format.
  
  Alternatively, if 
  Subversion\footnote{\url{http://subversion.apache.org/}} is installed on your
  system, you can check out the sources directly from our SourceForge repository
  by executing the following command:\\\cmd{svn co 
  https://gate.svn.sourceforge.net/svnroot/gate/mimir/trunk\\mimir}.
  The URL used in the command is for the {\tt trunk} version; adjust accordingly
  to point to one of the tags, if you require a release version instead. This
  will create a new directory called {\tt mimir} containing all the source
  files.
  \item {\bf Build the library:} change to the top level directory where you
  unpacked the downloaded \Mimir{} sources. If you can see the {\tt mimir-core},
  {\tt mimir-client}, {\em etc.} directories, then you are in the correct
  directory. Execute the \cmd{ant} command. This will download all the
  required dependencies, compile all the \Mimir{} libraries, and build the {\tt
  mimir-web} Grails plugin.
  \item {\bf Generate the demo application:} to use \Mimir{} you need a web
  application that you can access from your browser. The simplest \Mimir{}-based
  web application is {\tt mimir-demo}, which can be automatically generated
  and built by executing the \cmd{ant mimir-demo} command. This will produce a
  new directory, named {\tt mimir-demo}, located under the top level \Mimir{}
  directory.
  \item {\bf Run the demo application:}  change to the newly created {\tt
  mimir-demo} directory (\cmd{cd mimir-demo}) and execute the \cmd{grails prod
  run-app} command. This will start the application and will notify you which
  URL you should use in your browser to access it (normally
  \url{http://localhost:8080/mimir-demo/}).
\end{enumerate}
%
\section{Create, Populate, and Search an Index}
%
\begin{enumerate}
\setcounter{enumi}{4}
  \item {\bf Set-up your new \Mimir{} application:}
  navigate to the administration page. You will be prompted to configure your \Mimir{} instance.
  After clicking the link, enter the path to a local writable directory
  where new indexes will be created, and click the {\em Update}.button.
  \item \label{step:create} {\bf Create a new index:} navigate back to the
  administration page (by clicking the link at the top of the page). Under the {\em Local Indexes}
  section, click the {\em create a new local index} link. Give it a name (e.g.
  {\tt `test'}), and click the {\em create} button. Back on the administration
  page, click the name of the newly created index. This will take you to the
  index details page, where you can find the {\em Index URL} attribute. Make a
  note of its value, as you will need it later.
  \item {\bf Populate the new index:} 
  \begin{enumerate}
    \item Start GATE Developer, load the ANNIE application (Main Menu 
    $\rightarrow$ File $\rightarrow$ Load ANNIE System $\rightarrow$ with
    Defaults).
    \item Open the CREOLE Plugin Manager ((Main Menu $\rightarrow$ File
    $\rightarrow$ Manage CREOLE Plugins), and add a new plugin directory
    pointing at the {\tt mimir-client} directory inside the \Mimir{}
    distribution. Make sure the new plugin is loaded by checking the
    appropriate check-box.
    \item Load a new instance of {\em \Mimir{} Indexing PR} (Main Menu
    $\rightarrow$ File $\rightarrow$ New Processing Resource $\rightarrow$
    \Mimir{} Indexing PR), and add it to the end of the ANNIE application.
    \item Make sure that the {\tt mimirIndexUrl} parameter for the new PR is set
    to the {\em Index URL} value obtained at Step~\ref{step:create}.
    \item Load some test documents (e.g. some web pages from news web sites),
    create a GATE Corpus, add all the documents to the corpus, and set the
    newly corpus as the target for the ANNIE application.
    \item Run the ANNIE application. This will annotate the documents created
    during the previous step. The \Mimir{} Indexing PR instance will make
    sure the annotated documents are sent for indexing to your new Local Index.
  \end{enumerate}
  \item Go back to the index details page in your browser. Click the {\em Close}
  button, and wait for the index to finish closing. When done, go back to the
  main administration page. 
  \item {\bf You can now search the new index} by clicking the {\em search} link
  next to the name of your new index.
\end{enumerate}

To shut down the running web application, create a file named {\tt
.kill-run-app} in the {\em mimir-demo} directory, and wait for the application
to shut itself down.

To deploy the \Mimir{} web application to an application server (such as Apache
Tomcat) run the \cmd{grails prod war} command in the mimir-demo directory. A
{\tt mimir-demo-\{version\}.war} file will be created for you in the {\tt
target} sub-directory.
